### 2020-04-28 Milestones/miletsone5.Rmd
organizing my repo

---
title: "GOV 1006: Milestone 5"
subtitle: "Paper Replication"
author: "Pieter Quinton"
date: "3/31/2020"
output: 
  bookdown::pdf_document2:
    keep_tex: true
bibliography: milestone5_bib.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(stargazer)
library(plyr)
library(tidyverse)
```

## Introduction

The paper I have chosen for my replication project is entitled "Unemployment, Trust in Government, and Satisfaction with Democracy: An Empirical Investigation" (@bauer2018). ^[The replication materials for this paper and my subsequent analyses can be found at my [Github Repo](https://github.com/PGQuinton/gov1006-finalproject-2.0)] It was authored by Paul C. Bauer in 2018. The purpose of the paper is to evaluate how unemployment affects a person's trust in the government and their views towards democracy in general. Bauer begins with macro level data from the Netherlands and Switzerland to construct his analysis but expands to utilize panel data so as to evaluate the relationship on an individual level.

The aim of Bauer's paper is to show how unemployment influences a person's trust in government, satisfaction with democracy, and life satisfaction in general. Bauer also controls for variations in age, education, and party membership. Because the link between life satisfaction and unemployment is somewhat more widely studied, Bauer uses that as a base metric to evaluate the magnitude of the other effects upon. All three dependent variables (trust in government, satisfaction in democracy, and life satisfaction) are measured on a 1-10 scale with 0 indicating no trust or satisfaction and 10 indicating complete trust or satisfaction.

Bauer finds that there is no consistent effect between unemployment and trust in the government or satisfaction with democracy. These findings contradict his initial hypothesis that unemployment would lead people to have lower levels of trust and satisfaction with the government and democratic institutions. However, Bauer does find that unemployment is negatively related to life satisfaction. People who are unemployed are generally less satisfied with their lives than those who are employed. This finding is consistent with previous research on the topic. Moving forward, Bauer hopes to increase the sample size of the studied group to further explore heterogenous treatment effects across various subgroups of individuals. For example, he hypothesizes that the effect of unemployment on attitudes may be greater among people traditionally disadvantaged in the labor market such as women and workers of a lower-class.

## Replication

In the below section, I make use of the data collected by Bauer to conduct an analysis on how unemployment, age, and education level affect a person's life satisfaction. I run three regressions, each adding an additional explanatory variable in the above ordering. This is a similar analysis to that conducted in Table 3 of his paper, though I only include the LISS data from The Netherlands. My regressions make use of several packages including the tidyverse, stats, and stargazer (@R-tidyverse; @R-stargazer). The inspiration for my analysis comes from the King et. al. article entitled "Making the Most of Statistical Analyses: Improving Interpretation and Presentation" (@king2000).

```{r, results='asis'}

# Read in the data from the LISS study and save it as data

data <- read_rds("data-long.rds")

# Select the variables I want for my regressions

x <- data %>%
  select(education, age, member, life_satisfaction, unemployed) %>%
  drop_na()

# Run three linear regressions, each time adding another explanatory variable

fit1 <- lm(life_satisfaction ~ unemployed, data = x)

fit2 <- lm(life_satisfaction ~ unemployed + age, data = x)

fit3 <- lm(life_satisfaction ~ unemployed + age + education, data = x)

# Display the results in a singular stargazer table
# Set header to false to hide the commentary from the creator
# Renamed the covariate labels and dependent variable to match the formatting in the paper
# Added a title and source note

stargazer(fit1, fit2, fit3,
          header = FALSE,
          covariate.labels = c("Unemployment", "Age", "Education"),
          dep.var.labels = "Life Satisfaction",
          title = "Effects of Unemployment on Life Satisfaction",
          notes = "Data from Bauer 2018"
          )
```

The results of my regressions confirm the general findings of Bauer. Like he, I also found that unemployment negatively effects a person's life satisfaction. I did not test the effects of unemployment on trust in government and satisfaction in democracy though I will do so in future analyses.

\newpage

## Appendix

```{r}

# Loaded the LISS data from the directory

load(file = "./data-long-liss.RData")

# Saved the loaded data as liss

liss <- data.long.liss

# Loaded the SHP data from the directory

load(file = "./data-long-shp.RData")

# Saved the data as SHP

shp <- data.long.shp
```

```{r, results='asis'}

# Set results = 'asis' to get the stargazer output to show up

# Reformatted the liss data to get it into the manner in which it is presented in the table
# Caluclated the total number of responses for each year broken down by employed vs unemployed
# Spread that data to make it wide rather than long

liss_1 <- liss %>%
  group_by(time, unemployed) %>%
  summarize(n = n()) %>%
  drop_na() %>%
  spread(time, n) %>%
  select(-unemployed)

# Did the same reformating for the shp data and then added the liss data beneath it using
# rbind_fill
# Changed all NA values to blanks to match the table formatting

shp_1 <- shp %>%
  group_by(time, unemployed) %>%
  summarize(n = n()) %>%
  drop_na() %>%
  spread(time, n) %>%
  select(-unemployed) %>%
  rbind.fill(liss_1) %>%
  replace(is.na(.), "")

# Added rownames to the conjoined table in accordance with what they are in the paper

row.names(shp_1) <- c("SHP: employed", "SHP: unemployed", "LISS: employed", "LISS: unemployed")

# Displayed the table using stargazer
# Set header to false to hide the comments from the creator of the package
# Set summary to false to just display the table as is
# Changed font size and column separation width to get the table the right size for the pdf

stargazer(
  shp_1,
  header = FALSE,
  summary = FALSE,
  font.size = "small",
  column.sep.width = "-7.5pt",
  title = "Unemployment across Years",
  notes = " LISS = Longitudinal Internet Studies for the Social Sciences; SHP = Swiss Household Panel."
)
```

**The actual table from the paper is below:**

```{r image, out.width="100%"}

# I wasn't sure how to add an image but I found this technique on stackoverflow and seems to
# work well.
# You can set the width of the image in the {r} section above which I did

knitr::include_graphics("milestone5.png")
```

This photo is displayed using a function from the knitr package (@R-knitr).
\newpage

## References