### 2020-05-07 analysis.Rmd
added captions to my tables plus some other paper edits

---
title: "Paper Analysis Code"
author: "Pieter Quinton"
date: "4/23/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(rsconnect)
library(stargazer)
library(plyr)
library(plm)
library(wfe)
library(berryFunctions)
library(kableExtra)
library(tidyverse)
```

```{r data}

# Loaded the LISS data from the directory

load(file = "./Dataverse/data-long-liss.RData")

# Saved the loaded data as liss

liss <- data.long.liss

# Loaded the SHP data from the directory

load(file = "./Dataverse/data-long-shp.RData")

# Saved the data as SHP

shp <- data.long.shp
```

```{r extension data}

liss <- liss %>%
  group_by(unit) %>%
  arrange(time) %>%
  mutate(lagged_unemployed = lag(unemployed)) %>%
  ungroup(unit)

shp <- shp %>%
  group_by(unit) %>%
  arrange(time) %>%
  mutate(lagged_unemployed = lag(unemployed)) %>%
  ungroup(unit)
```

```{r extension table1, results = 'asis'}

# SHP
# I run six regressions
# Two for each question (trust in government, satisfaction with democracy, and life satisfaction)
# The first regression is just with unemployed as the indenpendent variable
# The second regression includes the controls (age, education, and member)
# The regression uses a pooling technique (plm) to track observations across units (households) 
# over time

shp.fit.pooling.tig.ext <- plm(trust_in_government ~ unemployed * lagged_unemployed, 
                        data = shp,
                        index = c("unit", "time"),
                        model = "pooling")

shp.fit.pooling.swd.ext <- plm(satisfaction_with_democracy ~ unemployed * lagged_unemployed, 
                        data = shp,
                        index = c("unit", "time"),
                        model = "pooling")

shp.fit.pooling.ls.ext <- plm(life_satisfaction ~ unemployed * lagged_unemployed, 
                        data = shp,
                        index = c("unit", "time"),
                        model = "pooling")

shp.fit.pooling.control.tig.ext <- plm(trust_in_government ~ unemployed * lagged_unemployed + age + education  + member, 
                        data = shp,
                        index = c("unit", "time"),
                        model = "pooling")

shp.fit.pooling.control.swd.ext <- plm(satisfaction_with_democracy ~ unemployed * lagged_unemployed + age + education  + member, 
                        data = shp,
                        index = c("unit", "time"),
                        model = "pooling")

shp.fit.pooling.control.ls.ext <- plm(life_satisfaction ~ unemployed * lagged_unemployed + age + education  + member, 
                        data = shp,
                        index = c("unit", "time"),
                        model = "pooling")

# LISS
# Same as above but with the LISS data instead of the SHP data

liss.fit.pooling.tig.ext <- plm(trust_in_government ~ unemployed * lagged_unemployed, 
                        data = liss,
                        index = c("unit", "time"),
                        model = "pooling")

liss.fit.pooling.swd.ext <- plm(satisfaction_with_democracy ~ unemployed * lagged_unemployed, 
                        data = liss,
                        index = c("unit", "time"),
                        model = "pooling")

liss.fit.pooling.ls.ext <- plm(life_satisfaction ~ unemployed * lagged_unemployed, 
                        data = liss,
                        index = c("unit", "time"),
                        model = "pooling")

liss.fit.pooling.control.tig.ext <- plm(trust_in_government ~ unemployed * lagged_unemployed + age + education  + member, 
                        data = liss,
                        index = c("unit", "time"),
                        model = "pooling")

liss.fit.pooling.control.swd.ext <- plm(satisfaction_with_democracy ~ unemployed * lagged_unemployed + age + education  + member, 
                        data = liss,
                        index = c("unit", "time"),
                        model = "pooling")

liss.fit.pooling.control.ls.ext <- plm(life_satisfaction ~ unemployed * lagged_unemployed + age + education  + member, 
                        data = liss,
                        index = c("unit", "time"),
                        model = "pooling")

# I display the results using a stargazer table
# I needed to omit the results in the footer like adj.rsq because they were preventing me from 
# adjusting the separation width between columns
# I can't figure out how to add a split header above the dependent variables to separate between 
# SHP and LISS
# I will keep working on this

stargazer(shp.fit.pooling.tig.ext, shp.fit.pooling.control.tig.ext, shp.fit.pooling.swd.ext, shp.fit.pooling.control.swd.ext, shp.fit.pooling.ls.ext, shp.fit.pooling.control.ls.ext, liss.fit.pooling.tig.ext, liss.fit.pooling.control.tig.ext, liss.fit.pooling.swd.ext, liss.fit.pooling.control.swd.ext, liss.fit.pooling.ls.ext, liss.fit.pooling.control.ls.ext,
          type = "latex",
          table.placement = "H",
          out="./extension-table-1.tex",
          omit.stat=c("LL","ser","f","adj.rsq"),
          header = FALSE,
          font.size = "tiny",
          digits = 2,
          column.sep.width = "-10pt",
          title = "SHP Switzerland and LISS Netherlands: pooled models with lagged unemployment",
          column.labels = c("(M1)", "(M2)", "(M3)", "(M4)", "(M5)", "(M6)", "(M7)", "(M8)", "(M9)", "(M10)", "(M11)", "(M12)"), 
          covariate.labels = c("Unemployed", "Lagged Unemployed", "UE:Lagged UE", "Age", "Education", "Member"),
          dep.var.labels = c("trust in government", "statisfaction with democracy", "life satisfaction", "trust in government", "satisfaction with democracy", "life satisfaction"),
          model.names = FALSE,
          model.numbers = FALSE,
          order = c(1,2,6,3,4,5),
          notes = "\\parbox[t]{15cm}{The above table presents the results from a series of pooled effects regressions examining the relationship between unemployment and trust in government, satisfaction with democracy, and life satisfaction. The regression includes an interaction between unemployment (present year) and lagged unemployment (previous year) to test for the possibility of hetergenous effects between short and long-term unemployment. Columns 1-6 are regression outputs from the SHP survey data while columns 7-12 are regression outputs from the LISS survey data. Looking at the results, the 'Unemployed' coefficients are statistically across all questions for both surveys with and without controls. However, the lagged unemployment coefficient is only statistically significant for the SHP data and the interaction coefficient is only significant for the life satisfaction question, for both surveys. }"
          )
```

```{r extension table2, results = 'asis'}

# SHP
# The same process as above but I used a fixed effects model instead of a pooled model
# The only difference is I change model from "pooled" to "within"

shp.fit.fe.tig.ext <- plm(trust_in_government ~ unemployed * lagged_unemployed, 
                        data = shp,
                        index = c("unit", "time"),
                        model = "within")

shp.fit.fe.swd.ext <- plm(satisfaction_with_democracy ~ unemployed * lagged_unemployed, 
                        data = shp,
                        index = c("unit", "time"),
                        model = "within")

shp.fit.fe.ls.ext <- plm(life_satisfaction ~ unemployed * lagged_unemployed, 
                        data = shp,
                        index = c("unit", "time"),
                        model = "within")

shp.fit.fe.control.tig.ext <- plm(trust_in_government ~ unemployed * lagged_unemployed + age + education  + member, 
                        data = shp,
                        index = c("unit", "time"),
                        model = "within")

shp.fit.fe.control.swd.ext <- plm(satisfaction_with_democracy ~ unemployed * lagged_unemployed + age + education  + member, 
                        data = shp,
                        index = c("unit", "time"),
                        model = "within")

shp.fit.fe.control.ls.ext <- plm(life_satisfaction ~ unemployed * lagged_unemployed + age + education  + member, 
                        data = shp,
                        index = c("unit", "time"),
                        model = "within")

# LISS
# Same as for SHP but with the LISS data

liss.fit.fe.tig.ext <- plm(trust_in_government ~ unemployed * lagged_unemployed, 
                        data = liss,
                        index = c("unit", "time"),
                        model = "within")

liss.fit.fe.swd.ext <- plm(satisfaction_with_democracy ~ unemployed * lagged_unemployed, 
                        data = liss,
                        index = c("unit", "time"),
                        model = "within")

liss.fit.fe.ls.ext <- plm(life_satisfaction ~ unemployed * lagged_unemployed, 
                        data = liss,
                        index = c("unit", "time"),
                        model = "within")

liss.fit.fe.control.tig.ext <- plm(trust_in_government ~ unemployed * lagged_unemployed + age + education  + member, 
                        data = liss,
                        index = c("unit", "time"),
                        model = "within")

liss.fit.fe.control.swd.ext <- plm(satisfaction_with_democracy ~ unemployed * lagged_unemployed + age + education  + member, 
                        data = liss,
                        index = c("unit", "time"),
                        model = "within")

liss.fit.fe.control.ls.ext <- plm(life_satisfaction ~ unemployed * lagged_unemployed + age + education  + member, 
                        data = liss,
                        index = c("unit", "time"),
                        model = "within")

# Stargazer output with the same formatting as Table 3

stargazer(shp.fit.fe.tig.ext, shp.fit.fe.control.tig.ext, shp.fit.fe.swd.ext, shp.fit.fe.control.swd.ext, shp.fit.fe.ls.ext, shp.fit.fe.control.ls.ext, liss.fit.fe.tig.ext, liss.fit.fe.control.tig.ext, liss.fit.fe.swd.ext, liss.fit.fe.control.swd.ext, liss.fit.fe.ls.ext, liss.fit.fe.control.ls.ext,
          type = "latex",
          out="./extension-table-2.tex",
          omit.stat=c("LL","ser","f","adj.rsq"),
          header = FALSE,
          font.size = "tiny",
          digits = 2,
          label = "table-2",
          column.sep.width = "-10pt",
          title = "SHP Switzerland and LISS Netherlands: fixed effects models with lagged unemployment",
          column.labels = c("(M13)", "(M14)", "(M15)", "(M16)", "(M17)", "(M18)", "(M19)", "(M20)", "(M21)", "(M22)", "(M23)", "(M24)"),
          covariate.labels = c("Unemployed", "Lagged Unemployed", "UE:Lagged UE", "Age", "Education", "Member"),
          dep.var.labels = c("trust in government", "statisfaction with democracy", "life satisfaction", "trust in government", "satisfaction with democracy", "life satisfaction"),
          model.names = FALSE,
          model.numbers = FALSE,
          order = c(1,2,6,3,4,5),
          notes = "\\parbox[t]{15cm}{The above table presents the results from a series of fixed effects regressions examining the relationship between unemployment and trust in government, satisfaction with democracy, and life satisfaction. The regression includes an interaction between unemployment (present year) and lagged unemployment (previous year) to test for the possibility of hetergenous effects between short and long-term unemployment. Columns 13-18 are regression outputs from the SHP survey data while columns 19-24 are regression outputs from the LISS survey data. Looking at the results, the 'Unemployed' coefficients are statistically significant for the life satisfaction question for both surveys with and without controls. However, the lagged unemployment coefficient is only statistically significant for the satisfaction with democracy question on the SHP survey while the interaction coefficient is significant for the satisfaction with democracy on the LISS survey.}"
          )
```

```{r extension table3 data, include = FALSE}

# SHP
# Same regressions but this time with the weighted fixed effects
# I didn't really understand how these were created so I largely mimicked what Bauer did in his
# replication code

shp.fit.wfe.tig.ext <- wfe(trust_in_government ~ unemployed * lagged_unemployed, 
                        data = shp,
                        treat = "unemployed",
                       unit.index = "unit", time.index = "time", method = "unit",
               qoi = "ate", estimator =NULL, hetero.se=TRUE, auto.se=TRUE,
               White = TRUE, White.alpha = 0.05, verbose = TRUE)

shp.fit.wfe.swd.ext <- wfe(satisfaction_with_democracy ~ unemployed * lagged_unemployed, 
                        data = shp,
                        treat = "unemployed",
                       unit.index = "unit", time.index = "time", method = "unit",
               qoi = "ate", estimator =NULL, hetero.se=TRUE, auto.se=TRUE,
               White = TRUE, White.alpha = 0.05, verbose = TRUE)

shp.fit.wfe.ls.ext <- wfe(life_satisfaction ~ unemployed * lagged_unemployed, 
                        data = shp,
                        treat = "unemployed",
                      unit.index = "unit", time.index = "time", method = "unit",
               qoi = "ate", estimator =NULL, hetero.se=TRUE, auto.se=TRUE,
               White = TRUE, White.alpha = 0.05, verbose = TRUE)

shp.fit.wfe.control.tig.ext <- wfe(trust_in_government ~ unemployed * lagged_unemployed + age + education  + member, 
                        data = shp,
                        treat = "unemployed",
                        unit.index = "unit", time.index = "time", method = "unit",
               qoi = "ate", estimator =NULL, hetero.se=TRUE, auto.se=TRUE,
               White = TRUE, White.alpha = 0.05, verbose = TRUE)

shp.fit.wfe.control.swd.ext <- wfe(satisfaction_with_democracy ~ unemployed * lagged_unemployed + age + education  + member, 
                        data = shp,
                        treat = "unemployed",
                        unit.index = "unit", time.index = "time", method = "unit",
               qoi = "ate", estimator =NULL, hetero.se=TRUE, auto.se=TRUE,
               White = TRUE, White.alpha = 0.05, verbose = TRUE)

shp.fit.wfe.control.ls.ext <- wfe(life_satisfaction ~ unemployed * lagged_unemployed + age + education  + member, 
                        data = shp,
                        treat = "unemployed",
                        unit.index = "unit", time.index = "time", method = "unit",
               qoi = "ate", estimator =NULL, hetero.se=TRUE, auto.se=TRUE,
               White = TRUE, White.alpha = 0.05, verbose = TRUE)

# LISS
# Same as SHP but with LISS data

liss.fit.wfe.tig.ext <- wfe(trust_in_government ~ unemployed * lagged_unemployed, 
                        data = liss,
                        treat = "unemployed",
                        unit.index = "unit", time.index = "time", method = "unit",
               qoi = "ate", estimator =NULL, hetero.se=TRUE, auto.se=TRUE,
               White = TRUE, White.alpha = 0.05, verbose = TRUE)

liss.fit.wfe.swd.ext <- wfe(satisfaction_with_democracy ~ unemployed * lagged_unemployed, 
                        data = liss,
                        treat = "unemployed",
                        unit.index = "unit", time.index = "time", method = "unit",
               qoi = "ate", estimator =NULL, hetero.se=TRUE, auto.se=TRUE,
               White = TRUE, White.alpha = 0.05, verbose = TRUE)

liss.fit.wfe.ls.ext <- wfe(life_satisfaction ~ unemployed * lagged_unemployed, 
                        data = liss,
                        treat = "unemployed",
                       unit.index = "unit", time.index = "time", method = "unit",
               qoi = "ate", estimator =NULL, hetero.se=TRUE, auto.se=TRUE,
               White = TRUE, White.alpha = 0.05, verbose = TRUE)

liss.fit.wfe.control.tig.ext <- wfe(trust_in_government ~ unemployed * lagged_unemployed + age + education  + member, 
                        data = liss,
                        treat = "unemployed",
                        unit.index = "unit", time.index = "time", method = "unit",
               qoi = "ate", estimator =NULL, hetero.se=TRUE, auto.se=TRUE,
               White = TRUE, White.alpha = 0.05, verbose = TRUE)

liss.fit.wfe.control.swd.ext <- wfe(satisfaction_with_democracy ~ unemployed * lagged_unemployed + age + education  + member, 
                        data = liss,
                        treat = "unemployed",
                        unit.index = "unit", time.index = "time", method = "unit",
               qoi = "ate", estimator =NULL, hetero.se=TRUE, auto.se=TRUE,
               White = TRUE, White.alpha = 0.05, verbose = TRUE)

liss.fit.wfe.control.ls.ext <- wfe(life_satisfaction ~ unemployed * lagged_unemployed + age + education  + member, 
                        data = liss,
                        treat = "unemployed",
                        unit.index = "unit", time.index = "time", method = "unit",
               qoi = "ate", estimator =NULL, hetero.se=TRUE, auto.se=TRUE,
               White = TRUE, White.alpha = 0.05, verbose = TRUE)

# The above regressions do not create an output that can be displayed through stargazer
# Instead I needed to manually run the regression with the included weights and then manually
# input those values into stargazer
# Again, I didn't really know how to do this so I just followed what Bauer did.

  regressors_int <- c("Unemployed", "Lagged")

  regressors_cont <- c("Age", "Education", "Member")

  d <- as.data.frame(matrix(rnorm(10 * 7), nc = 7))

  dep.var1 <- "Trust_in_government"

  names(d) <- c(dep.var1, regressors_int, regressors_cont)

  f <- as.formula(paste(dep.var1, "~ 0 +", paste(regressors_int, collapse = "*"), "+", paste(regressors_cont, collapse = "+")))

  p1 <- lm(f, d)

  dep.var2 <- "Satisfaction_with_democracy"

  names(d) <- c(dep.var2, regressors_int, regressors_cont)

   f <- as.formula(paste(dep.var2, "~ 0 +", paste(regressors_int, collapse = "*"), "+", paste(regressors_cont, collapse = "+")))

  p2 <- lm(f, d)

  dep.var3 <- "Life_satisfaction"

  names(d) <- c(dep.var3, regressors_int, regressors_cont)

   f <- as.formula(paste(dep.var3, "~ 0 +", paste(regressors_int, collapse = "*"), "+", paste(regressors_cont, collapse = "+")))

  p3 <- lm(f, d)
```

```{r stargazer data}
wfe_regression_data <- data.frame(
  c1 = as.numeric(summary(shp.fit.wfe.tig.ext)[2]$coefficients[,1]),
  c2 = as.numeric(summary(shp.fit.wfe.swd.ext)[2]$coefficients[,1]),
  c3 = as.numeric(summary(shp.fit.wfe.ls.ext)[2]$coefficients[,1]),
  c4 = as.numeric(summary(liss.fit.wfe.tig.ext)[2]$coefficients[,1]),
  c5 = as.numeric(summary(liss.fit.wfe.swd.ext)[2]$coefficients[,1]),
  c6 = as.numeric(summary(liss.fit.wfe.ls.ext)[2]$coefficients[,1]),
  se1 = as.numeric(summary(shp.fit.wfe.tig.ext)[2]$coefficients[,2]),
  se2 = as.numeric(summary(shp.fit.wfe.swd.ext)[2]$coefficients[,2]),
  se3 = as.numeric(summary(shp.fit.wfe.ls.ext)[2]$coefficients[,2]),
  se4 = as.numeric(summary(liss.fit.wfe.tig.ext)[2]$coefficients[,2]),
  se5 = as.numeric(summary(liss.fit.wfe.swd.ext)[2]$coefficients[,2]),
  se6 = as.numeric(summary(liss.fit.wfe.ls.ext)[2]$coefficients[,2]),
  t1 = as.numeric(summary(shp.fit.wfe.tig.ext)[2]$coefficients[,3]),
  t2 = as.numeric(summary(shp.fit.wfe.swd.ext)[2]$coefficients[,3]),
  t3 = as.numeric(summary(shp.fit.wfe.ls.ext)[2]$coefficients[,3]),
  t4 = as.numeric(summary(liss.fit.wfe.tig.ext)[2]$coefficients[,3]),
  t5 = as.numeric(summary(liss.fit.wfe.swd.ext)[2]$coefficients[,3]),
  t6 = as.numeric(summary(liss.fit.wfe.ls.ext)[2]$coefficients[,3]),
  p1 = as.numeric(summary(shp.fit.wfe.tig.ext)[2]$coefficients[,4]),
  p2 = as.numeric(summary(shp.fit.wfe.swd.ext)[2]$coefficients[,4]),
  p3 = as.numeric(summary(shp.fit.wfe.ls.ext)[2]$coefficients[,4]),
  p4 = as.numeric(summary(liss.fit.wfe.tig.ext)[2]$coefficients[,4]),
  p5 = as.numeric(summary(liss.fit.wfe.swd.ext)[2]$coefficients[,4]),
  p6 = as.numeric(summary(liss.fit.wfe.ls.ext)[2]$coefficients[,4])
) %>%
  insertRows(3:5)
```

```{r extension table3 output, results = 'asis'}

# Stargazer output
# I manually pull out the coefficient values from the regressions and add those to the
# skeleton of the regression that was created above
# I have to do this for each coefficient, each standard error, and each t and p value

# Then I format the stargazer table in the same manner as the previous two

  stargazer(p1, p1, p2, p2, p3, p3, p1, p1, p2, p2, p3, p3,
    coef = list(wfe_regression_data$c1,
                as.numeric(summary(shp.fit.wfe.control.tig.ext)[2]$coefficients[,1]),
                wfe_regression_data$c2,
                as.numeric(summary(shp.fit.wfe.control.swd.ext)[2]$coefficients[,1]),
                wfe_regression_data$c3,
                as.numeric(summary(shp.fit.wfe.control.ls.ext)[2]$coefficients[,1]),

                wfe_regression_data$c4,
                as.numeric(summary(liss.fit.wfe.control.tig.ext)[2]$coefficients[,1]),
                wfe_regression_data$c5,
                as.numeric(summary(liss.fit.wfe.control.swd.ext)[2]$coefficients[,1]),
                wfe_regression_data$c6,
                as.numeric(summary(liss.fit.wfe.control.ls.ext)[2]$coefficients[,1])

                ),
    se = list(wfe_regression_data$se1,
                as.numeric(summary(shp.fit.wfe.control.tig.ext)[2]$coefficients[,2]),
                wfe_regression_data$se2,
                as.numeric(summary(shp.fit.wfe.control.swd.ext)[2]$coefficients[,2]),
                wfe_regression_data$se3,
                as.numeric(summary(shp.fit.wfe.control.ls.ext)[2]$coefficients[,2]),

                wfe_regression_data$se4,
                as.numeric(summary(liss.fit.wfe.control.tig.ext)[2]$coefficients[,2]),
                wfe_regression_data$se5,
                as.numeric(summary(liss.fit.wfe.control.swd.ext)[2]$coefficients[,2]),
                wfe_regression_data$se6,
                as.numeric(summary(liss.fit.wfe.control.ls.ext)[2]$coefficients[,2])
              ),
    t = list(wfe_regression_data$t1,
                as.numeric(summary(shp.fit.wfe.control.tig.ext)[2]$coefficients[,3]),
                wfe_regression_data$t2,
                as.numeric(summary(shp.fit.wfe.control.swd.ext)[2]$coefficients[,3]),
                wfe_regression_data$t3,
                as.numeric(summary(shp.fit.wfe.control.ls.ext)[2]$coefficients[,3]),

                wfe_regression_data$t4,
                as.numeric(summary(liss.fit.wfe.control.tig.ext)[2]$coefficients[,3]),
                wfe_regression_data$t5,
                as.numeric(summary(liss.fit.wfe.control.swd.ext)[2]$coefficients[,3]),
                wfe_regression_data$t6,
                as.numeric(summary(liss.fit.wfe.control.ls.ext)[2]$coefficients[,3])

             ),  
    p = list(wfe_regression_data$p1,
                as.numeric(summary(shp.fit.wfe.control.tig.ext)[2]$coefficients[,4]),
                wfe_regression_data$p2,
                as.numeric(summary(shp.fit.wfe.control.swd.ext)[2]$coefficients[,4]),
                wfe_regression_data$p3,
                as.numeric(summary(shp.fit.wfe.control.ls.ext)[2]$coefficients[,4]),

                wfe_regression_data$p4,
                as.numeric(summary(liss.fit.wfe.control.tig.ext)[2]$coefficients[,4]),
                wfe_regression_data$p5,
                as.numeric(summary(liss.fit.wfe.control.swd.ext)[2]$coefficients[,4]),
                wfe_regression_data$p6,
                as.numeric(summary(liss.fit.wfe.control.ls.ext)[2]$coefficients[,4])
             ),  
    type = "latex",
    out="./extension-table-3.tex",
    omit.stat = "all",
    notes = "\\parbox[t]{15cm}{The above table presents the results from a series of weighted pooled effects regressions examining the relationship between unemployment and trust in government, satisfaction with democracy, and life satisfaction. The regression includes an interaction between unemployment (present year) and lagged unemployment (previous year) to test for the possibility of hetergenous effects between short and long-term unemployment. Columns 25-30 are regression outputs from the SHP survey data while columns 31-36 are regression outputs from the LISS survey data. Looking at the results, the 'Unemployed' coefficient is statistically significant for the life satisfaction question on both surveys with and without controls. However, the lagged unemployment coefficient and the interaction coefficient are not statistically significant for any of the questions on either survey.}",
    font.size = "tiny",
    digits = 2,
    column.sep.width = "-10pt", 
    title = "SHP Switzerland and LISS Netherlands: weighted fixed effects models",
    digit.separator="",
    rownames = FALSE,
    column.labels = c("(M25)", "(M26)", "(M27)", "(M28)", "(M29)", "(M30)", "(M31)", "(M32)", "(M33)", "(M34)", "(M35)", "(M36)"), 
    model.names = FALSE,
    model.numbers = FALSE,
    star.cutoffs=c(0.05,0.01,0.001),
    covariate.labels = c("Unemployed", "Lagged Unemployed", "UE:Lagged UE", "Age", "Education", "Member"),
    dep.var.labels = c("trust in government", "statisfaction with democracy", "life satisfaction", "trust in government", "satisfaction with democracy", "life satisfaction"),
    header = FALSE,
    order = c(1,2,6,3,4,5)
    )  
```